Discussion

### 2018-08-28 doc/reviewer2_response.Rmd

---
title: "Reviewer 2 response"
output: 
  bookdown::word_document2:
    reference_docx: my_styles.docx
urlcolor: blue
link-citations: true
---

__The information that the authors added about the methods is useful for better understanding the design of the study. I still have unresolved questions about the statistical analysis and therefore the results. It is still not clear how the statistical analyses were performed with respect to colinear drivers. For example, in Figure 4, you show that pCO2, pH, and O2 all covary - so I interpret that to mean that if you detect a significant correlation of XX bioindicator with pCO2, you cannot distinguish that effect from a correlation with O2. If pCO2 and O2 covary, you cannot separate their effects and test for them independently, as shown in Figure 5. Furthermore, in your forum response, you say "Because the VIF value for carbonate chemistry parameters and oxygen were above the threshold, we have not attempted to separate the effects of carbonate chemistry parameters (Ωar, pH and/or pCO2) from oxygen." However, O2 and pH and pCO2 are all shown with independent p-values in Figure 5. So, it is not clear to me how you used the VIF results. Perhaps additional text to the methods is needed to describe what you did with the results of the VIFs and how that influenced the statistical analysis that resultsed in Figure 5. Perhaps you could include the list of VIF values to support your decisions.__

We appreciate the reviewer's concern regarding clarification of the methods and we offer an explanation here to better describe our approach.  The reviewer is absolutely correct about collinearity between some of the drivers and the inability to separate the effects, as for pCO2, ph, and O2.  Our previous statistical design, although insufficiently described, was developed specifically to emphasize these limitations and to account for these issues in both analysis and interpretation of results.  Our analysis design first explored general relationships among environmental and pteropod response measures using multivariate analysis (RDA) and simple pairwise correlations (Figs. 4 and 5).  The second portion of our analysis provided a more formal framework to identify combined effects of stressors using linear models of pairwise combinations of environmental variables to explain pteropod response (Figs. 6 and 7).  This second portion was informed by the preceding exploratory analyses and included restrictions on which variables were evaluated due to collinaerity, using VIF thresholds.  As such, the results from the second set of analyses were used directly to infer relationships becuase of a robust design that accounted for collinearity.  Figures 4 and 5 were intended to demonstrate how the parameters covaried and we did not use them to directly infer causation.

To address these concerns, we have added some content to the methods to better describe the methods and partciularly to distinguish between the different objectives of the first and second set of analyses (beginning on line 213, first paragraph of statistical analyses section):

"Several of the environmental variables were correlated, which presented a challenge for inferring causal relationships with pteropod response measures.  To address this issue, the analysis was conducted in two stages where the first was exploratory and the second was a more formal approach using a subset of the environmental variables to directly infer causation.  The first set of analyses included all variables to demonstrate the complexity of the environmental dataset and directly informed the second set of analyses, whereby variables that were associated with the response measures were further screened to reduce collinearity in subsequent models.  Details for both sets of analyses follow."

Finally, we note that the VIF values for all pairwise comparisons of collinearity of the environmental variables can be viewed in the supplementary online material indicated at the end of the mansucript.  Therefore, we have not included additional tables for this information.

### 2018-10-01 doc/reviewer2_response2.Rmd

---
title: "Reviewer 2 response 2"
output: 
  bookdown::word_document2:
    reference_docx: my_styles.docx
urlcolor: blue
link-citations: true
---

__The text addition to Methods section 2.4 is helpful, but it does not resolve my issue. I am specifically concerned about the statistical analyses behind Figure 5. I assume that these analyses were conducted using linear models for pairwise comparisons as described starting on p. 6, line 240. My argument remains that this is inappropriate, given that you found that some of the environmental variables were highly correlated in the exploratory RDA. Ultimately, the outcome may be similar and may not change your discussion and interpretation much, but I recommend redoing this part of the statistical analyses using collapsed variables that combine correlated environmental variables. For example, in the RDA, you found that carbonate chemistry parameters and dissolved oxygen are all highly correlated. Therefore, it is redundant and inappropriate to analyze their effects separately (separate rows and p-values in Figure 5). In that table, one variable that represents O2, pCO2, Omega_ar, and pH should presented (one row instead of four).__

__Furthermore, I think that latitude should be included in the RDA, as upwelling intensity generally increases with latitude in this system. Therefore, any correlations with abiotic varables and latitude should be identified and accounted for in the subsequent statistical analyses.__

We appreciate the concerns of the reviewer and we hope that our response here provides sufficient clarity on the intent of our analyses.  First and most importantly, we reiterate that the analysis in Figure 5 was presented to demonstrate exactly the issue that the reviewer notes and that the results jusified our conservative approach to screen collinear variables in the following analysis (i.e., those for Figs. 6 and 7).  Therefore, we feel there is value in presenting the correlations among all variables to emphasize the complexity of the system and to provide readers with an overview of associations that were more rigorously evaluated in the follow-up analysis.  The key distinction between figure 5 and the follow-up analyses is that we never claimed the results in figure 5 were causative.  Our conservative statistical design in the follow-up analyses provided us with a more solid foundation to infer mechanistic relationships.

We also emphasize that the text noted by the reviewer on p. 6, line 240 was not the analysis used for Figure 5.  The pairwise comparisons using linear models were the follow-up analyses that exluded collinear variables.  The stars in figure 5 are simply the significance of correlation coefficients (i.e., an association).

We feel there is still some disagreement about presenting correlation results in Figure 5 among multiple variables that co-vary, such as the parameters pH, pCO2, omega, and O2.  We recreated Figure 5 using the first principal component axis of a PCA of all environmental variables.  Nearly 100% of the variance was described by the first three axes: axis one explained the OA parameters pCO2, pH, O2, and Omega, axis 2 explained temperature and latitude, and axis 3 explained fluorescence.

![](prcompex_files/figure-html/pcaplo.png)
The first axis above was used in place of the four correlated OA parameters in the original Figure 5:

![](prcompex_files/figure-html/corploPC1.png)

Although this addresses the reviewer's concern, we feel this presentation complicates evaluation of the study system.  A reader would have to know which variables were described by PC1 and the approximate loadings.  We feel this presentation lacks the transparency provided by the original figure, nor does it change any of our conclusions.

Further, we want to emphasize the unique variation explained by latitude that was not associated with the OA parameters.  The loadings on the PC axes show a distinct separation of latitude on the second axis and the OA parameters on the first.  Therefore, we argue that latitude does not confound our interpration of pteropod response to acidification.

It is further noted that latitude is weakly associated with PC1 in the updated Figure 5, although this assocation is weak compared to the association with temperature.  Our PCA demonstrated that latitude is primarily associated with temperature and we did not include latitude in follow-up analyses because of no mechanistic interpretation on pteropod response.  We felt it unnecessary to include latitude in the RDA as a specific variable, but instead showed the points sized relative to latitude to emphasize the weak association.  We are confident that our exclusion of latitude from the analysis does not affect our conclusions.

### 2018-10-02 doc/prcompex.Rmd
no more lat in pc cor plot

---
output:
  html_document:
  keep_md: yes
self_contained: no
code_folding: hide
---

```{r, echo = F, warning = F, message = F}

knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(knitr)
library(gridExtra)
library(grid)
library(stargazer)
library(ggord)

@@

dat_cor <- ptedat %>% 
  select(one_of('CTD', phychr, biochr)) %>% 
  # na.omit %>% 
  left_join(envdat, by = 'CTD') %>% 
  data.frame %>% 
  remove_rownames %>% 
  left_join(PC1, by = 'CTD') %>% 
  column_to_rownames('CTD') %>% 
  select(-Ara, -pCO2, -O2, -pH)

@@

levs <- c("Fluor", "Lat", "PC1", "Temp", sort(biochr), sort(phychr))
labs <- c('Chla', 'Lat', 'PC1', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
prplo <- crs %>%
  separate(pr, c('cor', 'sig'), sep = ' ') %>%
  filter(var1 %in% levs & var2 %in% levs) %>%
  mutate(
    cor = as.numeric(cor),
    var1 = factor(var1, levels = rev(levs), labels = rev(labs)),
    var2 = factor(var2, levels = rev(levs), labels = rev(labs)),
    sig = gsub('ns', '', sig)
  )

@@

outlab <- data.frame(
  y = c(3.5, 10.5, 16.5),
  lab = c('Population/\nPhysiology', 'Cellular', 'Environment')
)

p <- ggplot(prplo) +
  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') +
  geom_text(aes(y = var1, x = var2, label = sig)) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 19, xmax = 19) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 19, xmax = 19) +
  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 19, xmax = 19) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 19.5, ymax = 19.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 19.5, ymax = 19.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 19.5, ymax = 19.5) +
  pbase +
  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0))) +
  geom_hline(yintercept = 6.5, size = 1.5) +
  geom_hline(yintercept = 14.5, size = 1.5) +
  geom_vline(xintercept = 6.5, size = 1.5) +
  geom_vline(xintercept = 14.5, size = 1.5)

>>>>

---
output:
  html_document:
    keep_md: yes
    self_contained: yes
code_folding: hide
---

```{r, echo = F, warning = F, message = F}
library(knitr)
opts_chunk$set(fig.align = 'center', message = F, echo = T, cache = F, dev = 'png', dev.args = list(family = 'serif'), dpi = 400, fig.pos = '!h', warning = F, background = 'white', out.width = '100%',
               fig.process = function(x) {
                 x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                 if (file.rename(x, x2)) x2 else x
               })

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(knitr)
library(gridExtra)
library(grid)
library(stargazer)
library(ggord)

@@

dat_cor <- ptedat %>% 
  select(one_of('CTD', phychr, biochr)) %>% 
  # na.omit %>% 
  left_join(envdat, by = 'CTD') %>% 
  data.frame %>% 
  remove_rownames %>% 
  left_join(PC1, by = 'CTD') %>% 
  column_to_rownames('CTD') %>% 
  select(-Ara, -pCO2, -O2, -pH, -Lat)

@@

levs <- c("Fluor", "PC1", "Temp", sort(biochr), sort(phychr))
labs <- c('Chla', 'PC1', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', ' LPX', 'ORAC', 'ORACvLPX', 'SOD', 'abundance', 'dissolution', 'length', 'scarring', 'typeII', 'typeIII')
prplo <- crs %>%
  separate(pr, c('cor', 'sig'), sep = ' ') %>%
  filter(var1 %in% levs & var2 %in% levs) %>%
  mutate(
    cor = as.numeric(cor),
    var1 = factor(var1, levels = rev(levs), labels = rev(labs)),
    var2 = factor(var2, levels = rev(levs), labels = rev(labs)),
    sig = gsub('ns', '', sig)
  )

@@

outlab <- data.frame(
  y = c(3.5, 10.5, 16),
  lab = c('Population/\nPhysiology', 'Cellular', 'Environment')
)

p <- ggplot(prplo) +
  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') +
  geom_text(aes(y = var1, x = var2, label = sig)) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 18, xmax = 18) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 18, xmax = 18) +
  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[3], ymax = outlab$y[3], xmin = 18, xmax = 18) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 18.5, ymax = 18.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 18.5, ymax = 18.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[3], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[3], xmax = outlab$y[3], ymin = 18.5, ymax = 18.5) +
  pbase +
  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0))) +
  geom_hline(yintercept = 6.5, size = 1.5) +
  geom_hline(yintercept = 14.5, size = 1.5) +
  geom_vline(xintercept = 6.5, size = 1.5) +
  geom_vline(xintercept = 14.5, size = 1.5)
